%!TEX root = ./main.tex
\section{Related Work}
\label{sec6}
%Explain the work that are related to your problem, and to your three contributions.
\todo{checking the sentense}

As we discussed before, our work is much related to the pioneering work of \emph{resugaring} in \cite{resugaring}. The idea of "tagging" and "reverse desugaring" is a clear explanation of "resugaring", but it becomes very complex when the RHS of the desugaring rule becomes complex. Our approach does not need reverse desugaring, and is both more powerful and efficient.
For hygienic resugaring, compared with the approach of using DAG to solve the variable binding problem in \cite{hygienic}, our approach of "lazy desugaring" can achieve natural hygiene within our core language.

\emph{Macros as multi-stage computations} \cite{multistage} is a work related to our lazy expansion of sugars. Some other research \cite{modularstaging} about multi-stage programming \cite{MSP} indicates that it is useful for implementing domain-specific languages. However, multi-stage programming is a metaprogramming method, which mainly works for run-time code generation and optimization. In contrast, our lazy resugaring approach treats sugars as part of a mixed language, rather than separating them by staging. Moreover, lazy desugaring gives us a chance to derive evaluation rules of sugars, which is an advantage over multi-stage programming.

Our work is related to the \emph{Galois slicing for imperative functional programs} \cite{slicing}, a work for dynamically analyzing functional programs during execution. The forward component of the Galois connection maps a partial input $x$ to the greatest partial output $y$ that can be computed from $x$; the backward component of the Galois connection maps a partial output $y$ to the least partial input $x$ from which we can compute $y$.
%Our approach used a similar idea on slicing expressions and processing on subexpressions.
This can also be considered as a bidirectional transformation \cite{bx,lens07} and the round-tripping between desugaring and resugaring in the existing approach. In contrast to these works, our resugaring approach is basically unidirectional. 


There is a long history of hygienic macro expansion\cite{hygienicmacro}, and a formal specific hygiene definition was given by specific the binding scopes of macros \cite{10.5555/1792878.1792884}. Another formal definition of the hygienic macro\cite{EssenceofHygiene} is based on nominal logic\cite{10.1007/s001650200016}. Instead of using desugaring rules to achieve hygiene, we use lazy desugaring with small core language to avoid hygienic problem in our approach.
%
%When the tracking in their notation can be easily done for sugar whose rules can be derived automatically.

Our implementation is built upon the PLT Redex \cite{SEwPR}, a semantics engineering tool, but it is possible to implement our approach with other semantics engineering tools such as those in \cite{dynsem,Ksemantic} which aim to test or verify the semantics of languages. Their methods of can be easily combined with our approach to implement more general rule derivation. \emph{Ziggurat} \cite{Ziggurat} is a semantics extension framework, also allowing defining new macros with semantics based on existing terms of a language. It should be useful for static analysis of macros.
%Instead of semantics based on core language, the reduction rules of sugar derived by our approach is independent of the core language, which may be more concise for static analysis.
